---
title: "Seascape NMDS"
output: html_notebook
---

This Rmd reads data table containing taxa densities collected during IMARPE fisheries surveys (Peru) and seascape class identified at each sampled site from monthly and 8-day seascapes at 4km pixel. NMDS plots are created with these data to evaluate seascape affinity by taxonomic groups. These tables are generated by the Matlab routine 'imarpe_survey.m'. 

```{r}
library(vegan)
library(dplyr)
```


```{r}
pc_month = read.csv("taxa_table_month.csv", header= TRUE)
pc_8d = read.csv("taxa_table_8d.csv", header= TRUE)

#make community matrix - extract columns with abundance information
community_month = pc_month[, 1:(ncol(pc_month)-2)]
df_month = as.data.frame(community_month)
df_month[rowSums(df_month[])>0,]

community_8d = pc_8d[, 1:(ncol(pc_8d)-2)]
df_8d = as.data.frame(community_8d)
df_8d[rowSums(df_8d[])>0,]

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
```{r}
## run the metaMDS command from the vegan package to generate an NMDS plot.
set.seed(123)
nmds_month = metaMDS(df_month, distance = "bray")
nmds_month
```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
